---
title: "Bipartite"
author: "Juan Rocha"
output:
    html_document:
      theme:
        bootswatch: cosmo
        code_font:
            google: Fira Code
      df_print: paged
      code_folding: hide
      toc: true
      toc_float:
        collapsed: true
        smooth_control: true
      toc_depth: 3
      fig_caption: true
      highlight: pygments
      self_contained: false
      lib_dir: libs
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(fs)
```
This notebook perform some basic bipartite netowrk modeling on the exporting / importing companies layer in the data. `03-bipartite` replicates the idea but for municipalities / companies exporting.

## Data

Subnational soy exports data from Brazil based on the Trase dataset
```{r}
fls <- fs::dir_ls(path = "~/Documents/Projects/DATA/Trase/BRAZIL_SOY_2.6.0_pc/")

fls |> str_subset(".csv")

dat <- read_csv(file = str_subset(fls, ".csv")) |> 
    janitor::clean_names()
```


```{r message = FALSE, warning = FALSE}
library(ergm)
library(ergm.count)
library(ggnetwork)
library(ggnewscale)
library(tictoc)
library(patchwork)
```

Now there is 2020 data on the lastest dataset

```{r}
net2018 <- dat |> 
    filter(year == 2020) |> #filter(exporter == "DOMESTIC CONSUMPTION")
    select(exporter, importer, soy_equivalent_tonnes) |> 
    group_by(exporter, importer) |> 
    summarize(soy = sum(soy_equivalent_tonnes)) |>
    mutate(log_soy = log10(soy)) |> 
    mutate(exporter = str_to_title(exporter) |> str_c("e", sep="_"),
           importer = str_to_title(importer) |> str_c("i", sep="_")) |> 
    #filter(exporter != importer) |> # there are 14 self-loops
    network( matrix.type = "edgelist", ignore.eval = TRUE, bipartite = TRUE,
            loops = FALSE)

net2018
```

Calculate proportion of links with commitments: the following calculations derive the proportion per node with respect to all its connections, not the proportion of pair-wise links. That could be another avenue to model it, as link weight instead of node attribute. 

Geography can also be included but not sure if country or municipality is best? Or a combination? Municipality for Brazil, country for importer company. The problem is if the same company exports from multiple places.


```{r}
exp_commit <- dat |> 
    filter(year == 2020) |> 
     mutate(exporter = str_to_title(exporter) |> str_c("e", sep="_"),
           importer = str_to_title(importer) |> str_c("i", sep="_")) |> 
    select(exporter, commitment = zero_deforestation_brazil_soy) |> 
    mutate(commitment = case_when(
        is.na(commitment) ~ "missing",
        TRUE ~ commitment
    )) |> 
    group_by(exporter) |> 
    add_count() |> # number of outlinks total
    ungroup() |> group_by(exporter, commitment) |> 
    add_count(name = "n_type") |> 
    summarize(prop = n_type / n) |> unique() |> 
    mutate(commitment = str_to_lower(commitment) |> 
               str_replace_all(" ", "_")) |> 
    pivot_wider(names_from = commitment, values_from = prop, values_fill = 0)

exp_commit |> arrange(desc(company_commitment))
```

```{r}
imp_commit <- dat |> 
    filter(year == 2020) |> 
     mutate(exporter = str_to_title(exporter) |> str_c("e", sep="_"),
           importer = str_to_title(importer) |> str_c("i", sep="_")) |> 
    select(importer, commitment = zero_deforestation_brazil_soy) |> 
    mutate(commitment = case_when(
        is.na(commitment) ~ "missing",
        TRUE ~ commitment
    )) |> 
    group_by(importer) |> 
    add_count() |> # number of outlinks total
    ungroup() |> group_by(importer, commitment) |> 
    add_count(name = "n_type") |> 
    summarize(prop = n_type / n) |> unique() |> 
    mutate(commitment = str_to_lower(commitment) |> 
               str_replace_all(" ", "_")) |> 
    pivot_wider(names_from = commitment, values_from = prop, values_fill = 0)

imp_commit |> arrange(desc(company_commitment))
```

```{r}

exp_commit <- exp_commit |> 
    rename(node = exporter) |> 
    select(-missing, -none)

imp_commit <- imp_commit |> 
    rename(node = importer) |> 
    select(-missing, -none)

df_commit <- full_join(exp_commit, imp_commit)

df_commit <- df_commit |> ungroup() 

# get the attributes table on the same order as nodes in the network
df_attr <- tibble(
    node = network.vertex.names(net2018)
)

df_attr <- left_join(df_attr, df_commit)

## Add them as node covariates: updated for a bipartite undirected network
net2018 %v% "soy_moratorium" <- df_attr$soy_moratorium
net2018 %v% "company_commitment" <- df_attr$company_commitment
net2018 %v% "degree" <- sna::degree(net2018, gmode = "graph")
net2018 %v% "bet" <- sna::betweenness(net2018, gmode = "graph", rescale = TRUE)

```

```{r}
a <- exp_commit |> 
    ggplot(aes(company_commitment, soy_moratorium)) +
    geom_point(alpha = 0.5) +
    labs(title = "Exporting companies") +
    theme_light()
b <- imp_commit |> 
    ggplot(aes(company_commitment, soy_moratorium)) +
    geom_point(alpha = 0.5) +
    labs(title = "Importing companies") +
    theme_light()

a+b
```

```{r}
tbls <- dat |> 
    select(exporter, year, zero_deforestation_brazil_soy) |> 
    table()

dat |> pull(zero_deforestation_brazil_soy) |> unique()

tbls[,,] |> as_tibble() |> 
    ggplot(aes(year, n)) + 
    geom_line(aes(group = exporter), show.legend = FALSE) +
    facet_wrap(~zero_deforestation_brazil_soy, ncol = 1, scales = "free_y")

```



```{r}
ggplot(
    ggnetwork(net2018, arrow.gap = 0.01), 
    aes(x = x, y = y, xend = xend, yend = yend)) +
    geom_edges(
        aes(color = soy), size = 0.2,
        arrow = arrow(length = unit(4, "pt"), type = "open")) +
    scale_color_viridis_c(
        direction = -1, option = "B", trans = "log1p", alpha = 0.4,
        guide = guide_colorbar(title.position = "top")) +
    new_scale_color() +
    geom_nodes(aes(size = degree, color = company_commitment), 
               alpha = 1) + 
    scico::scale_color_scico(palette = "batlow", 
               guide = guide_colorbar(title.position = "top")) +
    # new_scale_color() +
    # geom_nodes(aes(size = idegree, color = company_commitment), 
    #            alpha = 0.75) + 
    # scico::scale_color_scico(
    #     palette = "batlow",
    #     guide = guide_colorbar(title.position = "top")) +
    scale_size(guide = guide_legend(title.position = "top")) +
    theme_void() +
    theme(legend.position = "bottom")
        
```


### ERGMs bipartite

```{r}
f0 <- ergm(net2018~edges)

summary(f0)
```

```{r}
f1 <- ergm(
    net2018 ~ edges + diff("company_commitment") + 
        diff("soy_moratorium"))
summary(f1)
```

```{r}
f2 <- ergm(
    net2018 ~ edges + diff("company_commitment") + 
        diff("soy_moratorium") +
        nodecov("company_commitment") + 
        nodecov("soy_moratorium"))
summary(f2)
```
```{r}
f3 <- ergm(
    net2018 ~ edges + b1cov("company_commitment") +
        b2cov("company_commitment") +
        b1cov("soy_moratorium") +
        b2cov("soy_moratorium")
)

summary(f3)
```

```{r}
#Crashes
# tic()
# f4 <- ergm(
#     net2018 ~ nonzero + sum,
#     response = "log_soy",
#     reference = ~StdNormal)
# toc()
# summary(f4)
```



### Commitments

```{r}
exp_commit2 <- dat |> 
    filter(year == 2018) |> 
     mutate(exporter = str_to_title(exporter) |> str_c("e", sep="_"),
           importer = str_to_title(importer) |> str_c("i", sep="_")) |> 
    select(exporter, importer, commitment = zero_deforestation_brazil_soy) |> 
    mutate(commitment = case_when(
        is.na(commitment) ~ "missing",
        TRUE ~ commitment
    )) |> 
    group_by(exporter) |> 
    add_count() |> # number of outlinks total
    ungroup() |> group_by(exporter, commitment) |> 
    add_count(name = "n_type") |> 
    summarize(prop = n_type / n) |> unique() |> 
    mutate(commitment = str_to_lower(commitment) |> 
               str_replace_all(" ", "_")) |> 
    pivot_wider(names_from = commitment, values_from = prop, values_fill = 0)

exp_commit2 |> arrange(desc(company_commitment))
```

```{r}
imp_commit <- dat |> 
    filter(year == 2018) |> 
     mutate(exporter = str_to_title(exporter) |> str_c("e", sep="_"),
           importer = str_to_title(importer) |> str_c("i", sep="_")) |> 
    select(importer, commitment = zero_deforestation_brazil_soy) |> 
    mutate(commitment = case_when(
        is.na(commitment) ~ "missing",
        TRUE ~ commitment
    )) |> 
    group_by(importer) |> 
    add_count() |> # number of outlinks total
    ungroup() |> group_by(importer, commitment) |> 
    add_count(name = "n_type") |> 
    summarize(prop = n_type / n) |> unique() |> 
    mutate(commitment = str_to_lower(commitment) |> 
               str_replace_all(" ", "_")) |> 
    pivot_wider(names_from = commitment, values_from = prop, values_fill = 0)

imp_commit |> arrange(desc(company_commitment))
```


There are multiple transactions between pairs of companies (exporter, importer). Despite they are the same companies, not all transactions are labelled under the same certification scheme. I cannot aggregate per dyad because we introduce errors on the commitments variable.

```{r}
dat |>
    filter(year == 2018) |>
    group_by(exporter,importer) |>
    add_count(name = "N") |> 
    filter(zero_deforestation_brazil_soy == "COMPANY COMMITMENT") |>
    summarize(n = n(), prop = n/N) |> 
    ggplot(aes(prop)) + geom_density() + geom_vline(xintercept = 1, color = "orange")
```



Questions for Trase ppl:

- Who does the zero deforestation commitment? The exporting or importing company?: It seems it is the exporting company because they are the ones actually closer to the field, but the network graph does not support that.
- Domestic consumption exporter/importer? Any data on the companies?
- Why is the commitments variable an attribute of each transaction, not of the company?
- node attribute country?